---
title: "Water Pump EDA"
author: "Amritha"
date: "22 October 2018"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(corrplot)
library(ggplot2)
library(rpart)
library(dplyr)
library(caret)

library(lubridate)

wp_1<-read.csv("E://College stuff//SEM 3//Final project//Tanzania water pump//Training set labels.csv",na.strings = c("NA",""))
wp_2<-read.csv("E://College stuff//SEM 3//Final project//Tanzania water pump//Training set values.csv",na.strings = c("NA",""))

## Merge both dataframes

wp_train<-merge(wp_1,wp_2)

head(wp_train)
View((wp_train[1:5,]))
names(wp_train)
```


```{r}

class(wp_train$date_recorded)

wp_train$date_recorded<-as.character(wp_train$date_recorded)
wp_train$date_recorded<-as.Date(wp_train$date_recorded)
head(wp_train$date_recorded)
t<-ymd(wp_train$date_recorded)
class(t)

wp_train$year_recorded<-year(t)

```

## Distinguishing categorical and numeric columns

```{r}
d=function(x){
num=sapply(x, is.numeric)
fact=sapply(x,is.factor)
return (list(names(x)[fact],names(x)[num]))
}
d(wp_train)
```


```{r}
wp_train$amount_tsh[wp_train$amount_tsh==0]<-NA
wp_train$gps_height[wp_train$gps_height==0]<-NA
wp_train$construction_year[wp_train$construction_year==0]<-NA
wp_train$num_private<-NULL
wp_train$population[wp_train$population==0]<-NA


wp_train$funder[wp_train$funder==0]<-NA
wp_train$installer[wp_train$installer==0]<-NA
wp_train$district_code[wp_train$district_code==0]<-NA

#wp_train$construction_year<-as.factor(wp_train$construction_year)


```

## Summary of dataset

```{r}
summary(wp_train)
```




##  Percentage of missing values

```{r}
missing<-function(x){
  sapply(x,function(y)sum(is.na(y))/length(y)*100)
}
missing(wp_train)
```

### Percentage of outliers

```{r}
outlier<-function(x){
  num_col<-names(x)[sapply(x, is.numeric)]
  x<-x[,num_col]
  sapply(x,function(y)length(boxplot.stats(y)$out)/sum(!is.na(y))*100)
}
outlier(wp_train)

```

```{r}
summary(wp_train)
summary(wp_train$installer)
summary(wp_train$gps_height)

```



## Checking frequency of the values in the target group

```{r}
table(wp_train$status_group)
prop.table(table(wp_train$status_group))
```

## Checking proportion of status group against quantity

```{r}
table(wp_train$quantity,wp_train$status_group)
prop.table(table(wp_train$quantity,wp_train$status_group))

```


## Plotting Categorical variables

```{r}

colnames(wp_train)


ggplot(data=wp_train,aes(x=quantity_group,fill=status_group))+geom_bar()+theme_light()+ theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    theme(legend.position = "top")

ggplot(data=wp_train,aes(x=quality_group,fill=status_group))+geom_bar()+theme_light()+ theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    theme(legend.position = "top")

ggplot(data=wp_train,aes(x=extraction_type_class,fill=status_group))+geom_bar()+theme_light()+ theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    theme(legend.position = "top")

ggplot(data=wp_train,aes(x=source_type,fill=status_group))+geom_bar()+theme_light()+ theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    theme(legend.position = "top")

ggplot(data=wp_train,aes(x=waterpoint_type,fill=status_group))+geom_bar()+theme_light()+ theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    theme(legend.position = "top")

ggplot(data=wp_train,aes(x=payment,fill=status_group))+geom_bar()+theme_light()+ theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    theme(legend.position = "top")


ggplot(data=wp_train,aes(x=installer,fill=status_group))+geom_bar()+theme_light()+ theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    theme(legend.position = "top")

ggplot(data=wp_train,aes(x=management_group,fill=status_group))+geom_bar()+theme_light()+ theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    theme(legend.position = "top")

ggplot(data=wp_train,aes(x=permit,fill=status_group))+geom_bar()+theme_light()+ theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    theme(legend.position = "top")

ggplot(data=wp_train,aes(x=as.factor(year_recorded),fill=status_group))+geom_bar()+theme_light()+ theme(axis.text.x = element_text(angle=65, vjust=0.6)) + 
    theme(legend.position = "top")



#ggplot(data=wp_train,aes(x=construction_year,fill=status_group))+geom_bar()+theme_light()+ theme(axis.text.x = element_text(angle=65, vjust=0.6)) +  theme(legend.position = "top")


```

```{r}

```


## Plotting continuous variables

```{r}

boxplot(wp_train$amount_tsh)
boxplot(wp_train$population)
boxplot(wp_train$gps_height)

ggplot(data = wp_train,aes(x=as.factor(status_group),y=amount_tsh))+geom_boxplot()
ggplot(data = wp_train,aes(x=as.factor(status_group),y=population))+geom_boxplot()
ggplot(data = wp_train,aes(x=as.factor(status_group),y=gps_height))+geom_boxplot()

```

```{r}

ggplot(wp_train, aes(x = year_recorded)) + 
  geom_histogram(bins = 20) + 
  facet_grid( ~ status_group)+theme(axis.text.x = element_text(angle=75, vjust=0.6))


ggplot(subset(wp_train, construction_year > 0), aes(x = construction_year)) +
  geom_histogram(bins = 20) + 
  facet_grid( ~ status_group)+theme(axis.text.x = element_text(angle=75, vjust=0.6))

ggplot(wp_train, aes(x = amount_tsh)) + 
  geom_histogram(bins = 20) + 
  facet_grid( ~ status_group)+theme(axis.text.x = element_text(angle=75, vjust=0.6))

ggplot(subset(wp_train, amount_tsh > 0), aes(x = amount_tsh)) +
  geom_histogram(bins = 20) + 
  facet_grid( ~ status_group)+ scale_x_continuous(labels = scales::comma)+theme(axis.text.x = element_text(angle=75, vjust=0.6))

ggplot(wp_train, aes(x = gps_height)) + 
  geom_histogram(bins = 20) + 
  facet_grid( ~ status_group)

ggplot(subset(wp_train, gps_height > 0), aes(x = gps_height)) +
  geom_histogram(bins = 20) + 
  facet_grid( ~ status_group)+ scale_x_continuous(labels = scales::comma)+theme(axis.text.x = element_text(angle=75, vjust=0.6))

ggplot(subset(wp_train, gps_height > 500), aes(x = gps_height)) +
  geom_histogram(bins = 20) + 
  facet_grid( ~ status_group)+ scale_x_continuous(labels = scales::comma)+theme(axis.text.x = element_text(angle=75, vjust=0.6))


ggplot(wp_train, aes(x = population)) + 
  geom_histogram(bins = 10) + 
  facet_grid( ~ status_group)+theme(axis.text.x = element_text(angle=75, vjust=0.6))


ggplot(subset(wp_train, population < 9000), aes(x = population)) +
  geom_histogram(bins = 8) + 
  facet_grid( ~ status_group)+ scale_x_continuous(labels = scales::comma)+theme(axis.text.x = element_text(angle=75, vjust=0.6))




```

## 
creating correlation matrix for numerical columns

```{r}
num<-dplyr::select_if(wp_train, is.numeric)

C<-cor(na.omit(num))

corrplot(C,method="circle")
```

```{r}

## removing amount column since it has 70% missing values
wp_train$amount_tsh<-NULL

wp_train$gps_height[which(is.na(wp_train$gps_height))]<-mean(wp_train$gps_height,na.rm = T)

wp_train$population[which(is.na(wp_train$population))]<-median(wp_train$population,na.rm=T)




```


```{r}

# 

wp_train %>% select(funder) %>% group_by(funder) %>% summarise(count=n()) %>%  top_n(5)


wp_train$funder<-as.character(wp_train$funder)
wp_train$funder[which(is.na(wp_train$funder))]<-"Government of Tanzania"
wp_train$funder<-as.factor(wp_train$funder)

wp_train$installer
wp_train %>% select(installer) %>% group_by(installer) %>% summarise(count=n()) %>%  top_n(5)

wp_train %>% select(scheme_management) %>% group_by(scheme_management ) %>% summarise(count=n()) %>%  top_n(5)

wp_train %>% select(scheme_name) %>% group_by(scheme_name) %>% summarise(count=n()) %>%  top_n(5)

wp_train %>% select(permit) %>% group_by(permit) %>% summarise(count=n()) %>%  top_n(5)

wp_train %>% select(construction_year) %>% group_by(construction_year) %>% summarise(count=n()) %>%  top_n(5)

```

```{r}
wp_test<-read.csv("E://College stuff//SEM 3//Final project//Tanzania water pump//Test set values.csv")

```



```{r}
wp_test$amount_tsh[wp_test$amount_tsh==0]<-NA
wp_test$gps_height[wp_test$gps_height==0]<-NA
wp_test$construction_year[wp_test$construction_year==0]<-NA
wp_test$num_private<-NULL
wp_test$population[wp_test$population==0]<-NA


wp_test$funder[wp_test$funder==0]<-NA
wp_test$installer[wp_test$installer==0]<-NA
wp_test$district_code[wp_test$district_code==0]<-NA
#wp_test$construction_year[wp_test$construction_year==0]<-NA
#wp_test$construction_year<-as.factor(wp_test$construction_year)


```

```{r}

class(wp_test$date_recorded)

wp_test$date_recorded<-as.character(wp_test$date_recorded)
wp_test$date_recorded<-as.Date(wp_test$date_recorded)
head(wp_test$date_recorded)
t<-ymd(wp_test$date_recorded)
class(t)

wp_test$year_recorded<-year(t)

```

## Distinguishing categorical and numeric columns

```{r}
d=function(x){
  num=sapply(x, is.numeric)
  fact=sapply(x,is.factor)
  return (list(names(x)[fact],names(x)[num]))
}
d(wp_test)
```

## Summary of dataset

```{r}
summary(wp_test)
```




##  Percentage of missing values

```{r}
missing<-function(x){
  sapply(x,function(y)sum(is.na(y))/length(y)*100)
}
missing(wp_test)
```

### Percentage of outliers

```{r}
outlier<-function(x){
  num_col<-names(x)[sapply(x, is.numeric)]
  x<-x[,num_col]
  sapply(x,function(y)length(boxplot.stats(y)$out)/sum(!is.na(y))*100)
}
outlier(wp_test)

```

```{r}
summary(wp_test)
summary(wp_test$installer)
summary(wp_test$gps_height)

```



## Checking frequency of the values in the target group

## Plotting continuous variables

```{r}
boxplot(wp_test$amount_tsh)
boxplot(wp_test$population)
boxplot(wp_test$gps_height)

```



## 
creating correlation matrix for numerical columns

```{r}
num<-dplyr::select_if(wp_test, is.numeric)

C<-cor(na.omit(num))

corrplot(C,method="circle")
```

```{r}

## removing amount column since it has 70% missing values
wp_test$amount_tsh<-NULL

wp_test$gps_height[which(is.na(wp_test$gps_height))]<-mean(wp_test$gps_height,na.rm = T)

wp_test$population[which(is.na(wp_test$population))]<-median(wp_test$population,na.rm=T)

```

```{r}
## Decision tree
library(rpart)
model_tree <- rpart(as.factor(status_group)~quantity_group+quality_group+extraction_type_class+source_type+waterpoint_type+payment+management_group+year_recorded+gps_height+population,data = wp_train)

pred_dec_tree <- predict(model_tree, wp_train,type = "class")

wp_test$status_group<-NA
wp_test$status_group<-predict(model_tree,wp_test,type = "class")

head(wp_test)

head(wp_test$status_group)

submission1<-NA
submission1<-wp_test %>% select(id,status_group)

head(submission1)

write.csv(submission1,"E://College stuff//SEM 3//Final project//Tanzania water pump//submission1.csv",row.names=FALSE)

```

```{r}
library(randomForest)

mtry = round(sqrt(length(colnames(wp_train))-1))
model_rf <- randomForest(status_group~quantity_group+quality_group+extraction_type_class+source_type+waterpoint_type+payment+management_group+year_recorded+gps_height+population,data = wp_train,ntree = 300,mtry = mtry)

pred_model_rf<-predict(model_rf,wp_train)

wp_test$status_group<-NA
wp_test$status_group <- predict(model_rf,wp_test)


submission2<-wp_test %>% select(id,status_group)

head(submission2)

write.csv(submission2,"E://College stuff//SEM 3//Final project//Tanzania water pump//submission2.csv",row.names=FALSE)
```

```{r}
library(caret)

class(pred_dec_tree)
pred_dec_tree<-as.factor(pred_dec_tree)
# confuision matrix 1
confusionMatrix(pred_dec_tree, wp_train$status_group)

# confuision matrix 2
confusionMatrix(pred_model_rf, wp_train$status_group)


# variable importance 

importance(model_rf)

varImpPlot(model_rf)

```


```{r}

```




```{r}


# Observe the installer variable


wp_train$install_3 <- substr(tolower(wp_train$installer),1,3)
wp_train$install_3[wp_train$install_3 %in% c(" ", "", "0", "_", "-")] <- "other"

# Take the top 15 substrings from above by occurance frequency
install_top_15 <- names(summary(as.factor(wp_train$install_3)))[1:15]
wp_train$install_3[!(wp_train$install_3 %in% install_top_15)] <- "other"
wp_train$install_3 <- as.factor(wp_train$install_3)

# Table of the install_3 variable vs the status of the pumps
table(wp_train$install_3, wp_train$status_group)

# As row-wise proportions, install_3 vs status_group
prop.table(table(wp_train$install_3, wp_train$status_group), margin = 1)

# Create install_3 for the test set using same top 15 from above
wp_test$install_3 <- substr(tolower(wp_test$installer),1,3)
wp_test$install_3[wp_test$install_3 %in% c(" ", "", "0", "_", "-")] <- "other"
wp_test$install_3[!(wp_test$install_3 %in% install_top_15)] <- "other"
wp_test$install_3 <- as.factor(wp_test$install_3)

```

```{r}

set.seed(42)
# model_forest <- randomForest(as.factor(status_group) ~ longitude + latitude + extraction_type_group + quantity + waterpoint_type + construction_year + install_3,
#                              data = wp_train,
#                              ntree = 800, mtry=mtry)
# 

#set.seed(42)
# model_forest_1 <- randomForest(as.factor(status_group) ~ longitude + latitude + extraction_type_group + quality_group + quantity + waterpoint_type + construction_year + source_type  + water_quality + population + install_3, data = wp_train, ntree = 800, mtry=mtry)
# 
# 
# 
# # Predict using the training values
# pred_forest_train <- predict(model_forest, wp_train)
# 
# importance(model_forest)
# 
# confusionMatrix(pred_forest_train, train$status_group)
# 
# # Predict using the training values number 2
# pred_forest_train_1 <- predict(model_forest_1, wp_train)
# 
# importance(model_forest_1)
# 
# confusionMatrix(pred_forest_train_1, train$status_group)

# Predict using the test values
# pred_forest_test <- predict(model_forest, wp_test)
# 
# # Predict using the test values number 2
# pred_forest_test_1 <- predict(model_forest_1, wp_test)

model_forest_2 <- randomForest(status_group~quantity_group+longitude + latitude+quality_group+extraction_type_class+source_type+waterpoint_type+payment+management_group+year_recorded+gps_height+population+ source_type  + water_quality+install_3,data = wp_train,ntree = 700,mtry = mtry)

wp_test$status_group<-NA
wp_test$status_group<-predict(model_forest_2,wp_test)

pred_forest_test_2<-predict(model_forest_2,wp_test,type="class")

#confusionMatrix(pred_forest_test_2, wp_train$status_group)
submission_3<-wp_test %>% select(id,status_group)

#write.csv(submission_3, file = "E://College stuff//SEM 3//Final project//Tanzania water pump//submission_new3.csv",row.names=FALSE)

head(submission_3)

prop.table(table(wp_test$status_group))

```

```{r}

# Create submission data frame
#submission<-NA
#submission <-data.frame(test$id)
#test$status_group<-NA

#test$status_group<-predict(model_forest, test)

#submission<-test %>% select(id,status_group)

#submission$status_group <- pred_forest_test
#names(submission)[1] <- "id"
#names(submission)[2]<-"status_group"

#head(submission)
#View(submission)
#names(pred_forest_test)

# Create submission data frame
#submission_1<-NA
#submission_1 <- data.frame(test$id)
#submission$status_group <- pred_forest_test_1
#names(submission)[1] <- "id"
#names(submission)[2]<-"status_group"

#names(submission)

#test$status_group<-NA
#test$status_group<-predict(model_forest_1, test)

#submission_1<-test %>% select(id,status_group)

#head(submission_1)
# submission csv
#write.csv(submission, file = "E://College stuff//SEM 3//Final project//Tanzania water pump//submission_new1.csv",row.names=FALSE)

  # submission csv 2
#write.csv(submission_1, file = "E://College stuff//SEM 3//Final project//Tanzania water pump//submission_new2.csv",row.names=FALSE)

```



